\chapter*{Acknowledge}
The time in Boston College passed so fast. My google photo constantly reminds me of the life details six year ago, when I entered the US, purchasing furnitures in IKEA, moving around Boston with one day nowhere to go due to the empty day on the rental lease, settling down in Brighton, and starting my PhD journey at Boston College. As the graduation approaches, I have an intense feeling that there is a growing force pulling me back to the beginning of the journey. Similar to the instanton effects in a compact spacetime, fragments of memories flood in my mind, getting interferenced and becoming increasingly vivid. Now, standing at the end of the journey, I would like to take this opportunity to express my sincer gratitude to all the people who have helped me along the way.


I would like to express my deepest gratitude and respect to my advisor, Prof. Ying Ran. The open and extensive discussions in his office are the most joyful moments during my PhD life. Ying's style of research, asking right questions, and never taking unexamined points for granted, has largely shaped my academic personality. As an academic role model, Ying also led by example, teaching me important lessons on how to become an independent theorist. He always emphasized the importance of being honest with oneself: to be crystal clear on the physical picture, to slow down and think deeply and correctly. I would also emphasize Ying's kindness and gentleness. Ying was always welcoming of so-called ``stupid'' questions, yet he also encouraged me to think independently. Whenever I feel drained and frustrated, the talk with him would recharge and motivate me again. Ying also taught me many stills to be a physicist, not only computational techniques but also skills outside of physics. I still remember the day before my presentation on the the March Meeting, Ying spent the entire afternoon going through my slides, page by page, teaching me how to present in an engaging and attractive way. I am deeply grateful for Ying's guidence and extremely luck to be his student.


I would like to thank the the rest of my thesis committee: Prof. Kenneth Burch, Prof. Xiao Chen, and Prof. Qiong Ma. I thank Prof. Burch for many assistance during the early years, on both course selection and course exemption. That helps me to save time learning advanced courses. I thank Prof. Xiao Chen for being a friend bouldering at Central Rock Gym during the pandemic and for sharing his research in the field of quantum information. I appreciate Prof. Qiong Ma for introducing Moir\'e physics to our department, which directed my attention to this fascinating field.


I am deeply thankful to Prof. Jung Hoon Han, for the collaboration on SATHE projection and inspiring discussion during his sabbatical visit at Boston College, also for his help on my postdoc applications. I would like to thank Prof. Fazel Tafti for sharing his expertise in solid state physics during our collaborations on Kitaev materials and for his patience with my delayed calculations. I am also grateful to Prof. Liang Kong, for his hospitality during my visit to his group at SUSTech. Additionally, I extend my gratitude to the faculty members of the Physics Department, including Prof. Ziqiang Wang, Prof. Ilija Zeljkovic, Prof. David Broido, and Prof. P. Bakshi, for great experience with them as both a student and as a teaching assistant. My sincere thanks also go to the department staff, especially Jane Carter, Scott Bortolotto, and Nancy Chevry, whose kindness made my stay at Boston College a pleasant experience.


I also want to thank my colleagues and collaborators, both at Boston College and elsewhere. In particular, I would like to thank our senior group member Xu Yang. The condensed matter theory journal club we held for two years, with just the two of us as speakers, was highly beneficial. I gained much from his sharing and the questions he posed during my talks. I would also like to thank my roommates and close friends, Hong Li and Hanchen Liu. The countless discussions over delicious food in Allston, ranging from scanning tunneling microscope details to measurement-induced phase transitions, and from physics to politics, have broadened my perspective and reshaped my understanding of the world. I extend my gratitude to my academic friends, Zhihao Zhang, Yongxin Zeng, Jiaqi Cai, Junkai Dong, Chong Wang, and Prof. Yinghai Wu. The online discussions through Q\&A websites and WeChat have been incredibly helpful and inspiring for my research. I am also thankful to Prof. Subir Sachdev, Prof. Xiaogang Wen, and Prof. John McGreevy, for their open-access notes and slides, covering high-level knowledge to frontier research. For friends at BC, I would like to thank Yiping Wang and He Zhao for organizing numerous hiking and food exploration around Boston, refreshing my mind everytime I got frustrated. I would also like to thank Yiqiu Han, Owen Yang, Sylvia Chen, Boxi Song, Rule Yi, Xiaohan Yao, Shrinkhala Sharma, and my officemate Vikram Ravindranoth, Jayakrishnan M.P. Nair, Vincent Plisson, Kewen Huang, and Wenyao Liu. The happy moments spent with them prevented me from getting ``permanent head damage'' (PhD).



% \vspace{2em}
% \noindent May 2024\\
% Xiaodong Hu


\pagebreak


\begin{center}
    \vspace*{\fill}
    \huge\emph{To my parents}
    \vfill\null
\end{center}